ms_figures

### 2018-03-28 analysis/bai/trends_resilience_bai.Rmd
new update

# Get mean values by site and add sample deepth
s <- re %>% 
  group_by(site, disturb_year) %>% 
  summarise_at(c("rt", "rs", "rc", "rrs"), funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(d, by=c("disturb_year" = "year", "site"="site")) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear")) 
```

>>>>

# Get mean values by site and add sample deepth
s <- re %>% 
  group_by(site, disturb_year) %>% 
  summarise_at(c("rt", "rs", "rc", "rrs"), funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(d, by=c("disturb_year" = "year", "site"="site")) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

s %>% write.csv(file = here::here('/out/bai_resiliences_severes' ,'resiliences_bai_droughts.csv'), row.names = FALSE) 
```

### 2018-03-28 man/ms_figures.Rmd
new update

@@

**Figure 2.** EVI standardized anomaly during the period 2000-2016 for northern and southern populations. Error bars show standard error.

##### 
```{r, eval=FALSE, echo=FALSE}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

**Figure 3.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (*black fill circle*) with southern ones (*blue empty circle*). For BAI we compared northern population (San Juan, SJ; *black triangle*) with southerns populations: Cáñar-High (CA-High; *blue empty squares*) and Cáñar-Low (CA-Low; *blue fill squares*). Different letters above error bars indicate significant *post hoc* differences between groups (see material and methods).

@@

@@

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot. We only show chronologies with # trees > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) ($r_2$ = `r2`).

@@

**Figure 5.** Residual Tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three of most recent severe drought years (1995, 2005 and 2012).

##### 
```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

@@

**Figure 6.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site; and *dark blue* bars: high-elevation shouthern site. Asteriks indicate significant ($P < 0.05$) correlation coefficients.

@@

**Figure 7.** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

@@

**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis).

# Supplementary

**Appendix S1.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017 for 28 meteorological stations (from National Spanish Meteorological Services, AEMET) distributed around the Sierra Nevada area. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 (*dotted* line) and -2 (*dashed* line) standard deviation. *Blue* lines represent +1 (*dotted* line) and +2 (*dashed* line) standard deviation. We labelled all years with average values below -1SD. ***Inset plot***: cumulative precipitation during the hydrological years 2004-2005 (*blue line*) and 2011-2012 (*red line*). The boxplot representing the average from 1950-2015 period. Data from meteorological station Granada, Base Aérea (National Spanish Meteorological Services, AEMET).

@@

cp <- ggplot(df, aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth))) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Cumulative Precipitation (mm)')  + xlab('') + 
  # ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 4, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 4, colour = "red") + 
  annotate("text", label = "mean 1950-2015", x = 3, y = 600, size = 4, colour = "black") +
  theme(panel.grid = element_blank())
```

@@

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_text_repel(
    data = subset(presn, csum_mean < (meandf - 1*sddf)),
    aes(label = hyear)) +
  theme(panel.grid.minor = element_blank()) +
  ylab('Cumulative Precipitation (hydrological year)') + xlab('year')

@@

##### 
```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% 
  filter(d_duration > 2) %>%  
  mutate(year = ifelse(minyear == maxyear, 
                       minyear, 
                       paste0(minyear, '-', maxyear))) %>% 
  dplyr::select(-minyear, -maxyear, -cat, -month_peak) %>% 
  as.data.frame()

@@

**Appendix S3.** Drought events for Sierra Nevada based on SPEI index. A drought event starts in the month when SPEI falls below the threshold of -1.28 [@Pascoa2017]. A drought event is considered only when SPEI value are below threshold for at least two consecutive months [*e.g* @Spinoni2015; Spinoni2017a; @Pascoa2017]. The ***duration*** of a drought event is the number of consecutive months with the SPEI lower than a certain threshold. ***Severity*** of a drought event is the sum of the SPEI values (absolute values) during the duration of the drought event. ***Intensity*** and ***Lowest SPEI** refer to the mean and lowest value of SPEI respectively during the drought event duration.

@@

**Appendix S5.** Reilience vs. droughts

@@

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, tpe='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)

@@

>>>>

##### 
**Figure 2.** EVI standardized anomaly during the period 2000-2016 for northern and southern populations. Error bars show standard error.

@@

##### 
**Figure 3.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (*black fill circle*) with southern ones (*blue empty circle*). For BAI we compared northern population (San Juan, SJ; *black triangle*) with southerns populations: Cáñar-High (CA-High; *blue empty squares*) and Cáñar-Low (CA-Low; *blue fill squares*). Different letters above error bars indicate significant *post hoc* differences between groups (see material and methods).

```{r, eval=FALSE, echo=FALSE}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

@@

##### 
**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot. We only show chronologies with # trees > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) ($r_2$ = `r2`).

@@

##### 
**Figure 5.** Residual Tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three of most recent severe drought years (1995, 2005 and 2012).

@@

##### 
**Figure 6.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site; and *dark blue* bars: high-elevation shouthern site. Asteriks indicate significant ($P < 0.05$) correlation coefficients.

```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

@@

#####
**Figure 7.** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

@@

##### 
**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis).

@@

```
##### 
**Appendix S1.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 and -2 standard deviation (*dotted* and *dashed* lines respectively). *Blue* lines represent +1 and +2 standard deviation (*dotted* and *dashed* lines respectively). Years with average values below -1SD are labelled. Data from 28 meteorological stations distributed around Sierra Nevada area (from National Spanish Meteorological Services, AEMET). ***Inset plot***: cumulative precipitation during the hydrological years 2004-2005 (*blue line*) and 2011-2012 (*red line*). The boxplot representing the average from 1950-2015 period. Data from meteorological station Granada, Base Aérea.

@@

cp <- ggplot(df, aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth))) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1, size=7),
        axis.text.y = element_text(size=7),
        axis.title.y = element_text(size = 7)) + 
  ylab('Cumulative Precipitation (mm)')  + xlab('') + 
  # ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 2, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 2, colour = "red") + 
  annotate("text", label = "1950-2015", x = 3, y = 600, size = 2, colour = "black") +
  theme(panel.grid = element_blank())
```

@@

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() +
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_text_repel(
    data = subset(presn, csum_mean < (meandf - 1*sddf)),
    aes(label = hyear), 
    size = 3) +
  theme(panel.grid.minor = element_blank()) +
  ylab('Cumulative Precipitation (hydrological year)') + xlab('year')

@@

##### 
**Appendix S3.** Drought events for Sierra Nevada based on SPEI index. A drought event starts in the month when SPEI falls below the threshold of -1.28 [@Pascoa2017]. A drought event is considered only when SPEI value are below threshold for at least two consecutive months [*e.g* @Spinoni2015; Spinoni2017a; @Pascoa2017]. The ***duration*** of a drought event is the number of consecutive months with the SPEI lower than a certain threshold. ***Severity*** of a drought event is the sum of the SPEI values (absolute values) during the duration of the drought event. ***Intensity*** and ***Lowest SPEI** refer to the mean and lowest value of SPEI respectively during the drought event duration.

```{r}
# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% 
  filter(d_duration > 2) %>%  
  mutate(year = ifelse(minyear == maxyear, 
                       minyear, 
                       paste0(minyear, '-', maxyear))) %>% 
  dplyr::select(-minyear, -maxyear, -cat, -month_peak) %>% 
  as.data.frame()

@@

@@

##### 
**Appendix S5.** Resilience metrics of the tree-growth for the most severe drought events. *Left*: Resistance (*Rt*); *Center*: Recovery (*Rc*); *Right* Resilience (*Rs*). Points indicate average of resilience metrics for all populations. Error bar corresponds standard error. Resilience metrics were computed for each population (sample depth > 10) and drought event.

```{r, warning=FALSE, out.extra='angle=90', fig.height=5}
# see trends-resilience_bai.Rmd 
df <- read.csv(file=here::here('/out/bai_resiliences_severes' ,'resiliences_bai_droughts.csv'), header=TRUE)

#https://gist.github.com/ottadini/6882677 
lm_eqn = function(m) {
  # Displays regression line equation and R^2 value on plot
  # Usage:
  # p + annotate("text", x=25, y=300, label=lm_eqn(lm(y ~ x, df)), parse=TRUE)

  l <- list(#a = format(coef(m)[1], digits = 2),
            # b = format(abs(coef(m)[2]), digits = 2),
            r2 = format(summary(m)$r.squared, digits = 3));

  if (coef(m)[2] >= 0)  {
    # eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  } else {
    # eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l) 
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  }

  as.character(as.expression(eq));                 
}

sdf <- df %>% 
  filter(samps >= 10) %>% 
  group_by(disturb_year) %>% 
  summarise_at(c("rt_mean", "rs_mean", "rc_mean", "rrs_mean"), 
               funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

plotrel <- function(df, v, x, label, XLab, YLab, coordX_eq, coordY_eq, titulo){
  require(ggplot2)
  require(ggrepel)

  formulita <- as.formula(paste0(v, "_mean_mean ~", x))

  yerrbar <- aes_string(ymin = paste0(v, "_mean_mean - ",  v, "_mean_se"), 
                        ymax = paste0(v, "_mean_mean + ",  v, "_mean_se"))

  ggplot(df, aes_string(y=paste0(v, "_mean_mean"), x=x, label=label)) + 
    geom_smooth(method='lm', se=FALSE, colour="black") +
    geom_errorbar(mapping = yerrbar) + 
    geom_point(size=4, pch = 21, colour="black", fill="white") + 
    theme_bw() + 
    xlab(XLab) + ylab(YLab) + 
    geom_text_repel() + 
    annotate("text", x=coordX_eq, y = coordY_eq, 
             label=lm_eqn(lm(formulita, df)), parse=TRUE) +
    ggtitle(titulo) + 
    theme(plot.title = element_text(size = 14, 
                                    face = "bold.italic",
                                    margin = margin(b = -15)), 
          axis.title.y = element_blank())
  }

sevRs <- plotrel(sdf, v = "rs", x = 'd_severity', label = 'disturb_year',
        XLab = "Drought Severity", YLab = "Resilience (Rs) (3year window size)", 
        coordX_eq = 7, coordY_eq = 0.5, titulo = "Rs")

sevRt <- plotrel(sdf, v = "rt", x = 'd_severity', label = 'disturb_year',
        XLab = "Drought Severity", YLab = "Resistance (Rt) (3year window size)",
        coordX_eq = 7, coordY_eq = 0.5, titulo = "Rt")

sevRc <- plotrel(sdf, v = "rc", x = 'd_severity', label = 'disturb_year',
        XLab = "Drought Severity", YLab = "Recovery (Rc) (3year window size)",
        coordX_eq = 7, coordY_eq = 0.5, titulo = "Rc")

sevRt + sevRc + sevRs

@@

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, type='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)

@@

##### 
**Appendix S7.** Growth commparations
